---
title: "Tutorial"
output: 
  learnr::tutorial:
    allow_skip: false
    progressive: true
runtime: shiny_prerendered
description: >
  The practical part of the "How to {ggplot2}" course.
---

```{r setup, include=FALSE}
library(learnr)
library(ggplot2)
library(gapminder)
knitr::opts_chunk$set(echo = FALSE)
```

## Part 1: Our first plot

### From scratch

Try to reproduce our scatter plot of the `gapminder` dataset, where we put `gdpPercap` on the x- and `lifeExp` on the y-axis and added a smoothed line.

```{r firstplot, exercise = TRUE}
ggplot(data = ?, mapping = ?) + 
  ? + 
  ?
```

```{r firstplot-solution}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) + 
  geom_point() + 
  geom_smooth()
```

### A bit too linear

Per default, {ggplot2} uses a generalized additive model to derive the smoothed line, but other options exist. Try adding `method = "lm"` as an argument to `geom_smooth()` to use a linear model instead. Is the linear model adequate?

```{r lm, exercise = TRUE}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) + 
  geom_point() + 
  geom_smooth()
```

```{r lm-solution}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

### Order matters

Does anything happen if you put the `geom_smooth()` function before the `geom_point()` function?

```{r order, exercise = TRUE}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) + 
  geom_smooth(method = "lm") +
  geom_point()
```

### Improving the scale

Probably, we could improve the graphic if we transform the x-axis from a linear to a log scale. Do you remember how to do this? Additionally, polish the x-axis format to dollar values.

```{r log, exercise = TRUE}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

```{r log-solution}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  scale_x_log10(scales::dollar)
```

### Bringing color into the game

There is another popular aesthetic that maps a variable to a color. Try the following and understand what happens:

```{r color, exercise = TRUE}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp, color = continent)) + 
  geom_point()
```

Say we want to ink each point green. Why does the following not work? How to fix it?

```{r green, exercise = TRUE}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp, color = "green")) + 
  geom_point()
```

```{r green-solution}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) + 
  geom_point(color = "green")
```

### Too much is too much

Consider the following plot. We have a smoothing line for each separate continent. This might be a bit confusing, can you change the code such that we have only one overall smoothing line but still have the points color-coded by continents?

```{r one-smooth, exercise = TRUE}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp, color = continent, fill = continent)) + 
  geom_point() + 
  geom_smooth() +
  scale_x_log10()
```

```{r one-smooth-solution}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) + 
  geom_point(mapping = aes(color = continent)) + 
  geom_smooth() +
  scale_x_log10()
```

### Quiz time!

```{r quiz}
quiz(
  question("Which of the R packages listed below can we load to access the `ggplot()` function?",
    answer("gapminder"),
    answer("ggplot"),
    answer("ggplot2", correct = TRUE),
    answer("tidyverse", correct = TRUE)
  ),
  question("What is the `aes()` function good for? Telling {ggplot2} ...",
    answer("what data we use."),
    answer("what relationships we want to see.", correct = TRUE),
    answer("the plot type we want to produce."),
    answer("that we want to adjust the x-axis.")
  )
)
```

## Part 2: Facets and more geoms

### Everything on the line

Remember that we discussed the following code? Adjust it such that all facets are plotted in one row. You can consult `?facet_grip` for help.

```{r one-row, exercise = TRUE}
ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp)) +
  geom_line(color = "grey", aes(group = country)) +
  geom_smooth() +
  facet_wrap(~continent) +
  labs(x = "Year",
       y = "Life expectancy",
       title = "Life expectancy over time on five continents")
```

```{r one-row-solution}
ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp)) +
  geom_line(color = "grey", aes(group = country)) +
  geom_smooth() +
  facet_wrap(~continent, nrow = 1) +
  labs(x = "Year",
       y = "Life expectancy",
       title = "Life expectancy over time on five continents")
```

### A new data set

We looked at a new data set. Glimpse at it!

```{r data-overview, exercise = TRUE}
pillar::glimpse(socviz::gss_sm)
```

### Bar plots

We created a bar plot for religious preferences and I announced that we can transform the y-axis to show relative frequencies. The following does the trick. Try to remove the `group = 1` argument, is it required?

```{r barplot, exercise = TRUE}
ggplot(socviz::gss_sm,aes(x = religion)) +
  geom_bar(aes(y = ..prop.., group = 1))
```

Let's make it colorful! But somehow, the following is not as colorful as it could be. Where is the mistake? Expert level question: The legend is redundant, because the religions are already printed on the x-axis - how to remove the legend?

```{r barplot-color, exercise = TRUE}
ggplot(socviz::gss_sm,aes(x = religion, color = religion)) +
  geom_bar()
```

```{r barplot-color-solution}
ggplot(socviz::gss_sm,aes(x = religion, fill = religion)) +
  geom_bar() +
  guides(fill = "none") # to remove the legend
```

### Histograms

Executing the following code will result in a message and a warning (both not the end of the world). Try to understand their meaning.

```{r, histogram}
ggplot(data = socviz::gss_sm, mapping = aes(x = age)) +
  geom_histogram()
```

```{r, histogram-solution}
# the data set includes NA's which causes the warning
sum(is.na(socviz::gss_sm$age)) 

# choosing a number of bins silences the message
ggplot(data = socviz::gss_sm, mapping = aes(x = age)) +
  geom_histogram(bins = 20)
```

### Density plots

We used the following code to plot a density of the life expectations in the gapminder data set for the year 2007. Employ aesthetics to separate the densities by continent.

```{r, density}
ggplot(data = subset(gapminder::gapminder, subset = year == 2007), 
       mapping = aes(x = lifeExp)) +
  geom_density()
```

```{r, density-solution}
ggplot(data = subset(gapminder::gapminder, subset = year == 2007), 
       mapping = aes(x = lifeExp, fill = continent, color = continent)) +
  geom_density(alpha = 0.5)
```

All densities look (more or less) bimodal - any idea why that could be?

### Boxplots

Together we built the following boxplot. Exchange `geom_boxplot()` by `geom_violin()` to create a violin plot instead. Which one do you prefer?

```{r, boxplot}
ggplot(data = subset(gapminder::gapminder, subset = year == 2007), 
       mapping = aes(x = pop,
                     y = reorder(continent, pop))) +
  geom_boxplot() +
  scale_x_log10() + 
  labs(y = NULL,
       x = "Populations in 2007")
```

```{r, boxplot}
ggplot(data = subset(gapminder::gapminder, subset = year == 2007), 
       mapping = aes(x = pop,
                     y = reorder(continent, pop))) +
  geom_violin() +
  scale_x_log10() + 
  labs(y = NULL,
       x = "Populations in 2007")
```

### One question at the end...

```{r quiz}
question("When using `geom_bar()`, {ggplot2}...",
  answer("plots the raw data."),
  answer("transforms the data before plotting.", correct = TRUE)
)
```
