---
title: "Best Practices"
author: "Lennart Oelschl&auml;ger"
date: 'Last rendered on `r format(Sys.Date(),"%d.%m.%Y")`'
output: 
  slidy_presentation:
    smart: true
---

```{r, include = FALSE}
Sys.setenv(LANG = "en")
```
  
## Why and what 

Welcome to this tiny course on best practices in R! `r emo::ji("wave")`

### Why do we care?

...

### What is this course about?

...

### What do you need?

Basic R skills + a not-too-old version of R (>= 4.0.0) + RStudio

### At the end of the day...

...

### Sources

...

### Found mistakes? Have suggestions?

I'm sure you have! Please leave a note [here](https://github.com/loelschlaeger/rcourse/issues/new/choose). `r emo::ji("wish")`

## First topic

...

## Now it's your turn

This course includes practicals! `r emo::ji("strong")`

Executing the following lines gives you access to the course material:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("loelschlaeger/rcourse")
library(rcourse)
```

To start the practicals, type:

```{r, eval = FALSE}
practicals()
```

To open a copy of these slides, type:

```{r, eval = FALSE}
slides()
```


