---
title: "Good Practices"
author: "Lennart Oelschl&auml;ger"
date: 'Last rendered on `r format(Sys.Date(),"%d.%m.%Y")`'
output: 
  slidy_presentation:
    smart: true
---

```{r, include = FALSE}
Sys.setenv(LANG = "en")
```
  
## Why and what 

Welcome to this tiny course on good practices in R! `r emo::ji("wave")`

### Why do we care?

...

### What is this course about?

...

### What do you need?

Basic R skills + a not-too-old version of R (>= 4.0.0) + RStudio

### At the end of the day...

...

### Sources

...

### Found mistakes? Have suggestions?

I'm sure you have! Please leave a note [here](https://github.com/loelschlaeger/rcourse/issues/new/choose). `r emo::ji("wish")`

### Course material

Executing the following lines in R gives you access to the course material:

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("loelschlaeger/rcourse")
library("rcourse")
```

To open a copy of these slides, type:

```{r, eval = FALSE}
slides()
```

To start the practicals, type:

```{r, eval = FALSE}
practicals()
```

## Bad practices

Your colleague Bob gives you his code file named `stuff.r`:

```r
# Author: me
# Purpose: some data work

# here I get our data
data1<-"C:\Users\bob\MyDocuments\RProject\data\metrics.rds"
data2<- "C:\Users\bob\MyDocuments\RProject\data\prices.rda"
hannoverrückaktie="C:\Users\bob\MyDocuments\RProject\data\stocks.csv"

mean <- function(x) sum(x)

print(data1["symbol"])
print(data1["period"])
print(data1["inventory"])
print(data1["liabilities"])
print(mean(data1[3,]))

x = bobs_function(hannoverrückaktie)

bobs_function <- function(x) {}
```

## Code style

- Reproducible code (seed)
- Paths
- consistent
- 80 character lines
- `<-` instead of `=`

## Functions

- don't repeat yourself
- warning, error, tryCatch

## Debugging

- divide and conquer
- minimal reprex

## Version control

## R packages

## Markdown
