---
title: "Good Practices"
author: "Lennart Oelschl&auml;ger"
date: 'Last rendered on `r format(Sys.Date(),"%d.%m.%Y")`'
output: 
  slidy_presentation:
    smart: true
    highlight: rstudio
---

```{r, include = FALSE}
Sys.setenv(LANG = "en")
```
  
## Why and what 

Welcome to this tiny course on good practices in R! `r emo::ji("wave")`

### Why do we care?

...

### What is this course about?

...

### What do you need?

Basic R skills + a not-too-old version of R (>= 4.0.0) + RStudio

### At the end of the day...

...

### Sources

...

### Found mistakes? Have suggestions?

I'm sure you have! Please leave a note [here](https://github.com/loelschlaeger/rcourse/issues/new/choose). `r emo::ji("wish")`

### Course material

Executing the following lines in R gives you access to the course material:

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("loelschlaeger/rcourse")
library("rcourse")
```

To open a copy of these slides, type:

```{r, eval = FALSE}
slides()
```

To start the practicals, type:

```{r, eval = FALSE}
practicals()
```

## Bad practices

Your colleague Bob hands you a code file named `stuff.r` and expresses his desire to collaborate with you on his new project. Without any time for explanations, he simply requests, "please check out my code, we can discuss it later." `r emo::ji("boy")`

```r
# Author: me
# Purpose: finance data work

# here I get our data
prices="C:\Users\bob\MyDocuments\RProject\data\stocks.csv"
börsenaktienpreisevonmcdonalds <- prices %>% filter(symbol == "MCD") %>% mutate("open" = lag(close))

### Calculate Returns
return <- function(x) { (x$close - x$open) / x$open } 
börsenaktienpreisevonmcdonalds$rtn <- return(börsenaktienpreisevonmcdonalds)

# calculate moving average
bobs_function <- function(d, w) {
  ma <- NULL
  for (i in 1:(length(d) - w + 1)) ma <- c(ma, mean(d[i:(i + w - 1)]))
  ma
}
stk_MA10 <- bobs_function(börsenaktienpreisevonmcdonalds$close, 10)
börsenaktienpreisevonmcdonalds$MA10 <- c(rep(NA, 9), stk_MA10)

# do we have buy signal here?
buy_signal <- function(p, ma) if (p > ma) 'Buy' else 'Sell'
for (i in 10:1762) {
  if (buy_signal(börsenaktienpreisevonmcdonalds[i, "close"], börsenaktienpreisevonmcdonalds[i, "MA10"]) == "Buy") {
    börsenaktienpreisevonmcdonalds[i, "x"] <- buy_signal(börsenaktienpreisevonmcdonalds[i, "close"], börsenaktienpreisevonmcdonalds[i, "MA10"])
  }
}

write.csv(börsenaktienpreisevonmcdonalds, 'output.csv')
```

Is there anything you like / dislike about Bob's code? `r emo::ji("thumbsup")` `r emo::ji("thumbsdown")`

## Good practices

Following good coding practices helps to maintain consistency, readability, and understandability of the code. In the R programming language, adhering to a consistent code style is particularly important due to the language's flexibility and wide use in data analysis and visualization.

- Naming: Use meaningful and descriptive names for variables, functions, and data objects. In R, it is common to use `lower_case_with_underscores`, `period.separated.names`, or `camelCase`. 

- Indentation: Use consistent indentation to clearly delineate code blocks and improve readability.

- Spacing: Use spaces around operators, after commas, and between arguments in functions to make the code more readable.

- Commenting: Include comments to explain the purpose of code blocks, functions, and complex calculations. This helps other programmers and your future self understand the code more easily.

- Assignments: R allows both `<-` and `=` for assignment, but it's generally recommended to use `<-` for assignment and `=` for passing arguments in functions. Consistency helps avoid confusion and errors.

- Organizing code: Organize your code into logical sections or functions, separating data preparation, analysis, and visualization. This makes your code easier to understand, maintain, and debug.

- Following a style guide: To maintain consistency across projects and teams, it's helpful to follow a style guide. The Google R Style Guide and the tidyverse style guide are popular choices in the R community.

- Use of libraries: When using third-party libraries, follow the conventions and style used by those libraries. This will make your code more consistent and easier to understand for others who are familiar with those libraries.

- Version control: Use version control tools like Git to track changes and collaborate with others, allowing you to easily revert to previous versions of your code if needed.

- Reproducible code: By ensuring your code in R is reproducible, you contribute to the scientific integrity, transparency, and overall quality of your work, which is essential in research and data-driven fields. In R, you can simply use `set.seed()`.

- Paths: Use relative but never absolute paths for file inputs and outputs to ensure that the code can be run on different systems without modification.

- Presentation of results: Consider RMarkdown to create dynamic documents that combine code, results, and narrative, making it easier for others to understand and reproduce your work.

- Line length limit: Shorter lines are easier to read and understand, as they require less horizontal scanning by the reader. Long lines can be difficult to follow, increasing the cognitive load and the risk of misinterpretation. 80 character lines are recommended.

## Functions

- don't repeat yourself
- warning, error, tryCatch

## Debugging

- divide and conquer
- minimal reprex

## Version control and collaboration

## R packages

