---
title: "How to {ggplot2}"
author: "Lennart Oelschl&auml;ger"
date: '`r format(Sys.Date(),"%d.%m.%Y")`'
output: slidy_presentation
---
  
## Why and what 

Welcome to this tiny course on data visualization in R with [{ggplot2}](https://ggplot2.tidyverse.org/)! `r emo::ji("wave")`

### Why do we care?

Potentially, plots can [beautifully inform](https://en.wikipedia.org/wiki/Charles_Joseph_Minard#The_map_of_Napoleon's_Russian_campaign) or [horribly mislead](https://en.wikipedia.org/wiki/Misleading_graph#Truncated_graph). [Colors](https://michaelbach.de/ot) and [shape](http://vis.stanford.edu/files/2010-MTurk-CHI.pdf) matter! `r emo::ji("balance_scale")`

### Why {ggplot2}?

The {ggplot2} package implements a *grammar of graphics*, a series of distinct tasks to make a graphic.

### What is this course about?

Being in "decent control" of {ggplot2} to produce "meaningful" plots. 

### What do you need?

Basic R skills + a not-too-old version of R + RStudio

### At the end of the day...

### My sources

- Healy (2019): [Data Visualization: A practical introduction](https://socviz.co/)
- Wickham (2016): [ggplot2: elegant graphics for data analysis](https://ggplot2-book.org/)
- Wilkinson (2005): The Grammar of Graphics
- [Tufte](https://en.wikipedia.org/wiki/Edward_Tufte) (1983): The Visual Display of Quantitative Information

### Found mistakes? Have suggestions?

I'm sure you have! Please leave a note [here](https://github.com/loelschlaeger/howtoggplot2/issues/new/choose). `r emo::ji("wish")`

## Our first plot

Load {ggplot2}.

```{r}
# install.packages(ggplot2)
library(ggplot2)
```

We need data, let's go with an excerpt from the famous [Gapminder dataset](https://www.gapminder.org/tools):

```{r}
# install.packages(gapminder)
library(gapminder)
head(gapminder)
str(gapminder)
```

First, we tell the `ggplot()` function what data we use and what variables we wish to see on each axis:

```{r}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) 
```

Something is missing ... `r emo::ji("thinking")` We need an additional layer, a `geom_*` function!

```{r}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) +
  geom_point()
```

There are more of them which we can simply add (literally add!):

```{r, message=FALSE}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p <- p + geom_point() + geom_smooth()
p
```

As a last polishing step for now, we improve the x-axis scale and the plot labels.

```{r, message=FALSE}
p + scale_x_log10(labels = scales::dollar) +
  labs(x = "GDP per capita",
       y = "Life expectancy in years",
       title = "Economic growth as an indicator for life expectancy",
       subtitle = "Data points are country-years",
       caption = "Source: Gapminder")
```

Finally, we can use the `ggsave()` function to save our plot:

```{r, eval = FALSE}
ggsave("some_descriptive_name.pdf")
```

### The {ggplot2} workflow

1. Call `ggplot()`
2. Set `data = ...`
3. Set `mapping = aes(...)`
4. Add one (or more) `geom_*()` functions
5. Adjust the scale and labels

## It's your turn

This course includes tutorials! `r emo::ji("sunglasses")`

Executing the following lines gives you access to the course material:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("loelschlaeger/howtoggplot2")
library(howtoggplot2)
```

To start the tutorial, type:

```{r, eval = FALSE}
tutorial()
```

To open a copy of these slides, type:

```{r, eval = FALSE}
slides()
```

To submit an issue on GitHub about this course, type:

```{r, eval = FALSE}
issue()
```

## Facets and more geoms

Our goal is to plot the trajectory of life expectancy over time for each country in the gapminder data. 

```{r}
ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp)) +
  geom_line()
```

We must not forget to group by country! `r emo::ji("light_bulb")`

```{r}
ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp)) +
  geom_line(aes(group = country))
```

But can you make sense of this mess? Luckily, we can group by continents:

```{r}
ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp)) +
  geom_line(aes(group = country)) +
  facet_wrap(~continent)
```

Better don't `facet_wrap(~country)`... Let's polish our plot with the things we already learned:

```{r, message = FALSE}
ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp)) +
  geom_line(color = "grey", aes(group = country)) +
  geom_smooth() +
  facet_wrap(~continent) +
  labs(x = "Year",
       y = "Life expectancy",
       title = "Life expectancy over time on five continents")
```

Notice that we supplied a formula to `facet_wrap`. This can be more advanced, for example (with `facet_grid`):

```{r, message = FALSE, warning = FALSE}
ggplot(data = socviz::gss_sm, mapping = aes(x = age, y = childs)) +
  geom_point(alpha = 0.2) +
  geom_smooth() +
  facet_grid(sex ~ race) +
  labs(x = "Age",
       y = "Childs",
       title = "Relationship between age and number of childs",
       subtitle = "Separated by sex (in rows) and race (in columns)")
```

As a last input for this part, we learn four new geoms.

### Bar plots

```{r}
ggplot(data = socviz::gss_sm, mapping = aes(x = religion)) +
  geom_bar()
```

Using relative instead of absolute counts on the y-axis is covered in the tutorials.

### Histograms

```{r}
ggplot(data = socviz::gss_sm, mapping = aes(x = age)) +
  geom_histogram()
```

We address the message and the warning in the tutorials.

### Density plots

```{r}
ggplot(data = subset(gapminder::gapminder, subset = year == 2007), 
       mapping = aes(x = lifeExp)) +
  geom_density()
```

### Boxplots

```{r}
ggplot(data = subset(gapminder::gapminder, subset = year == 2007), 
       mapping = aes(x = pop,
                     y = reorder(continent, pop))) +
  geom_boxplot() +
  scale_x_log10() + 
  labs(y = NULL,
       x = "Populations in 2007")
```

We look at a variant on the basic boxplot that {ggplot2} offers in the tutorials.

## Annotations

We can plot text annotations to plots via `geom_text()`:

```{r}
ggplot(data = socviz::elections_historic, 
       mapping = aes(x = popular_pct,
                     y = ec_pct,
                     label = winner_label)) +
  geom_point() +
  geom_text()
```

This is hard to read. Adjusting the position is possible, but it is fuzzy and not robust. The extension {ggrepel} is designed to do this task for us:

```{r}
ggplot(data = socviz::elections_historic, 
       mapping = aes(x = popular_pct,
                     y = ec_pct,
                     label = winner_label)) +
  geom_point() +
  ggrepel::geom_text_repel()
```

We can also annotate only selected points (outliers for example) like follows:

```{r}
ggplot(data = socviz::elections_historic, 
       mapping = aes(x = popular_pct,
                     y = ec_pct,
                     label = winner_label)) +
  geom_point() +
  ggrepel::geom_text_repel(
    data = subset(socviz::elections_historic, popular_pct < 0.5 & ec_pct > 0.5)
  ) + 
  geom_hline(yintercept = 0.5) +
  geom_vline(xintercept = 0.5)
```

And finally, we can annotate anywhere almost everything we wish via `annotate()`:

```{r}
ggplot(data = socviz::elections_historic, 
       mapping = aes(x = popular_pct,
                     y = ec_pct,
                     label = winner_label)) +
  geom_point() +
  geom_hline(yintercept = 0.5) +
  geom_vline(xintercept = 0.5) +
  annotate(geom = "rect", xmin = 0, xmax = 0.5, ymin = 0, ymax = 0.5, fill = "red", alpha = 0.2) +
  annotate(geom = "text", x = 0.25, y = 0.25, label = "Some text.")
```