---
title: "How to {ggplot2}"
author: "Lennart Oelschl&auml;ger"
date: '`r Sys.Date()`'
output: slidy_presentation
---
  
## Why and what 

Welcome to this tiny course on data visualization in R with [{ggplot2}](https://ggplot2.tidyverse.org/)! `r emo::ji("wave")`

### Why do we care?

Potentially, plots can [beautifully inform](https://en.wikipedia.org/wiki/Charles_Joseph_Minard#The_map_of_Napoleon's_Russian_campaign) or [horribly mislead](https://en.wikipedia.org/wiki/Misleading_graph#Truncated_graph). [Colors](https://michaelbach.de/ot) and [shape](http://vis.stanford.edu/files/2010-MTurk-CHI.pdf) matter! `r emo::ji("balance_scale")`

### Why {ggplot2}?

The {ggplot2} package implements a *grammar of graphics*, a series of distinct tasks to make a graphic.

### What is this course about?

Being in decent control of {ggplot2} to produce meaningful plots. 

### Requirements

Some R skills. A not too old version of R. Optionally RStudio.

### References

- Wickham 2016
- Tufte (1983): The Visual Display of Quantitative Information
- Wilkinson 2005

### Found Errors? Have suggestions?

I'm sure you have! Please leave a note [here](). `r emo::ji("wish")`

## Our first plot

Load {ggplot2}.

```{r}
library(ggplot2)
```

We need data, what about ...

```{r}
library(gapminder)
head(gapminder)
str(gapminder)
```

First, we tell the `ggplot()` function what data we are using and what variables we wish to see on each axis:

```{r}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) 
```

Something is missing ... `r emo::ji("thinking")` We need an additional layer, a `geom_*` function!

```{r}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) +
  geom_point()
```

### Let's recap

1. Call `ggplot()`
2. Set `data = ...`
3. Set `mapping = aes(...)`
4. Add `+ geom_point()`

## It's your turn

Text
