---
title: "Good Practices"
author: "Lennart Oelschl&auml;ger"
date: "Version of `r format(Sys.Date(), '%d.%m.%Y')`"
output: 
  slidy_presentation:
    smart: true
    highlight: rstudio
---

```{r, include = FALSE}
Sys.setenv(LANG = "en")
```

## Why and what 

This is just a test.

Welcome to this tiny course on good practices in R! `r emo::ji("wave")`

### Why do we care?

Good practices in R enhance readability, maintainability, and collaboration, making code more efficient and less prone to errors.

### What is this course about?

We will discuss several good practices in R, including code formatting, exception handling, debugging. 

### What do you need?

Basic R skills + a not-too-old version of R (>= 2.10) + RStudio

### At the end of the day...

... you will have improved your code style, writing readable and maintainable code, using functions effectively, leveraging version control, and creating reproducible analyses.

### Course material

Executing the following lines in R gives you access to the course material:

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("loelschlaeger/rcourse", upgrade = "never")
library("rcourse")
```

To open a copy of these slides, type:

```{r, eval = FALSE}
slides()
```

To start the practicals, type:

```{r, eval = FALSE}
practicals()
```

### Found mistakes? Have suggestions?

You can leave a note [here on GitHub](https://github.com/loelschlaeger/rcourse/issues/new/choose). `r emo::ji("wish")`

## Reproducible science

In science, we aim for: `r emo::ji("man_scientist")` `r emo::ji("woman_scientist")`

1. Knowledge

2. Reproducible results

### How to gain knowledge?

- Learning and understanding

- Exploring and discovering

- Synthesizing and applying

### Why should we work reproducible?

- Convince with your work (yourself, supervisor, reviewer, public)

- Simplify further research

- Write better publications and improve your reputation

### How to work reproducible?

- Documentation (answer the *how?* and *why?*)

- Version control

- Make sure that work can be executed again with identical results

### Our focus

As statisticians, our computers serve as our laboratories, and our R codes act as our test tubes. `r emo::ji("test_tube")` `r emo::ji("computer")`

How experienced are you with R? What would you like to learn today? Please give me a quick feedback.

```{r, echo = FALSE, eval = TRUE, fig.align = 'center'}
url <- "https://forms.gle/PSJGa8h1svMgEvQP7"
plot(qrcode::qr_code(url))
title(url, cex.main = 2, line = 2)
```

## Code formatting

Your colleague Bob hands you a code file named `stuff.R` and expresses his desire to collaborate with you on his new research project. Without any time for explanations, he simply requests, "please check out my code, we can discuss it later." `r emo::ji("boy")`

```{r, eval = FALSE}
# Author: me
# Purpose: finance data work

# here I get our data
prices=read.csv("C:\Users\bob\MyDocuments\RProject\data\stocks.csv")
bC6rsenaktienpreisevonmcdonalds <- prices %>% filter(symbol == "MCD") %>% mutate("open" = lag(close))

### Calculate Returns
return <- function(x) { (x$close - x$open) / x$open } 
bC6rsenaktienpreisevonmcdonalds$rtn <- return(bC6rsenaktienpreisevonmcdonalds)

# calculate moving average
bobs_function <- function(d, w) {
  ma <- NULL
  for (i in 1:(length(d) - w + 1)) ma <- c(ma, mean(d[i:(i + w - 1)]))
  ma
}
stk_MA10 <- bobs_function(bC6rsenaktienpreisevonmcdonalds$close, 10)
bC6rsenaktienpreisevonmcdonalds$MA10 <- c(rep(NA, 9), stk_MA10)

# do we have buy signal here?
buy_signal <- function(p, ma) if (p > ma) 'Buy' else 'Sell'
for (i in 10:1762) {
  if (buy_signal(bC6rsenaktienpreisevonmcdonalds[i, "close"], bC6rsenaktienpreisevonmcdonalds[i, "MA10"]) == "Buy") {
    bC6rsenaktienpreisevonmcdonalds[i, "x"] <- buy_signal(bC6rsenaktienpreisevonmcdonalds[i, "close"], bC6rsenaktienpreisevonmcdonalds[i, "MA10"])
  }
}

write.csv(bC6rsenaktienpreisevonmcdonalds, 'output.csv')
```

What are the things you like / dislike about Bob's code? `r emo::ji("thumbsup")` `r emo::ji("thumbsdown")`

## Functions

Suppose you have a dataset of sales transactions, and you want to calculate the total sales, the total number of transactions, and the average sales amount for each day. `r emo::ji("woman_office_worker")` `r emo::ji("chart")`

```{r}
# Sample data frame
sales_data <- data.frame(
  date = as.Date(c("2023-01-01", "2023-01-01", "2023-01-02", "2023-01-02", "2023-01-03")),
  amount = c(50, 30, 40, 20, 60)
)

# Day 1
day1_data <- subset(sales_data, date == "2023-01-01")
total_sales_day1 <- sum(day1_data$amount)
transaction_count_day1 <- nrow(day1_data)
average_sales_day1 <- total_sales_day1 / transaction_count_day1

# Day 2
day2_data <- subset(sales_data, date == "2023-01-02")
total_sales_day2 <- sum(day2_data$amount)
transaction_count_day2 <- nrow(day2_data)
average_sales_day2 <- total_sales_day2 / transaction_count_day2

# Day 3
day3_data <- subset(sales_data, date == "2023-01-03")
total_sales_day3 <- sum(day3_data$amount)
transaction_count_day3 <- nrow(day3_data)
average_sales_day3 <- total_sales_day3 / transaction_count_day3
```

Instead of writing repetitive code, you can use a function:

```{r, eval = TRUE}
# Function to calculate sales metrics
calculate_sales_metrics <- function(data, date) {
  day_data <- subset(data, date == as.Date(date))
  total_sales <- sum(day_data$amount)
  transaction_count <- nrow(day_data)
  average_sales <- total_sales / transaction_count
  metrics <- list(
    "total_sales" = total_sales, 
    "transaction_count" = transaction_count, 
    "average_sales" = average_sales
  )
  return(metrics)
}

# Calculate sales metrics for each day
sales_metrics_day1 <- calculate_sales_metrics(sales_data, "2023-01-01")
sales_metrics_day2 <- calculate_sales_metrics(sales_data, "2023-01-02")
sales_metrics_day3 <- calculate_sales_metrics(sales_data, "2023-01-03")
```

By using a function, you can **avoid code duplication** and make your code more **maintainable** and **scalable**.

## Exceptions

What happens when something goes wrong with your R code? `r emo::ji("cold_sweat")`

### Warnings

Basic R functions throw warnings if unexpected things happen:

```{r}
log(-1)
```

Include warnings in your own code to inform the user about bad things:

```{r}
boil_egg <- function(minutes) {
 if (minutes < 5) {
   warning("Too short!")
 } else if (minutes >= 5 && minutes <= 8) {
   cat("Yummy!")
 } else {
   warning("Too long!")
 }
}
```

```{r}
boil_egg(minutes = 5)
boil_egg(minutes = 4)
boil_egg(minutes = 20)
```

### Errors

Warnings do not stop execution, but errors do:

```{r, error = TRUE}
1 + "a"
```

Sometimes, error messages are confusing:

```{r, error = TRUE}
calculate_sales_metrics(data = data, date = "2023-01-01")
```

### Error recovery

Detect, handle, and recover from errors that occur during the execution of a program.

```{r, eval = TRUE}
# Define a function that uses try() to handle errors
safe_calculate_sales_metrics <- function(data, date) {
  result <- try(calculate_sales_metrics(data = data, date = date), silent = TRUE)
  if (inherits(result, "try-error")) {
    cat("Unable to compute sales metrics, please check your inputs.")
    return(NULL)
  }
  return(result)
}
```

```{r, eval = TRUE}
safe_calculate_sales_metrics(data = data, date = "2023-01-01")
safe_calculate_sales_metrics(data = sales_data, date = "2023-01-01")
```

What is the problem with `data`?

```{r}
head(data)
```

## Debugging

- divide and conquer
- Getting help: minimal reprex, Stack Overflow

## Code performance and efficiency

- vectorized
- pre-allocated

## Overview of good practices in R

Following good coding practices helps to maintain consistency, readability, and intelligibility of the code. In R, adhering to a consistent code style is particularly important due to the language's flexibility and wide use in data analysis and visualization.

- **Naming:** Treat variable, function, and data object names like a secret handshake between you and your future self: make them meaningful and descriptive, so when you look back at your code, you'll both know exactly what's going on without needing a decoder ring.

- **Spacing and indentation:** Always use consistent spacing and indentation to improve code readability.

- **Line limits:** Shorter lines are easier to read and understand, as they require less horizontal scanning by the reader. Long lines can be difficult to follow, increasing the cognitive load and the risk of misinterpretation.

- **Commenting:** Include comments to explain the purpose of code blocks, functions, and complex calculations. This helps other programmers and your future self understand the code more easily.

- **Assignments:** R allows both `<-` and `=` for assignment, but it's generally recommended to use `<-` for assignment and `=` for passing arguments in functions. 

- **Organizing code:** Organize your code into logical sections or functions, separating data preparation, analysis, and visualization. This makes your code easier to understand, maintain, and debug.

- **Following a style guide:** To maintain consistency across projects and teams, it's helpful to follow a style guide. The tidyverse style guide is a popular choice in the R community.

- **Reproducible code:** By ensuring your code in R is reproducible, you contribute to the scientific integrity, transparency, and overall quality of your work, which is essential in research and data-driven fields. In R, you can simply use `set.seed()`.

- **Paths:** Use relative but never absolute paths for file inputs and outputs to ensure that the code can be run on different systems without modification.

### Outlook

- **Version control:** Use version control tools like Git to track changes and collaborate with others, allowing you to easily revert to previous versions of your code if needed.

- **Presentation of results:** Consider RMarkdown to create dynamic documents that combine code, results, and narrative, making it easier for others to understand and reproduce your work.

- **Use the tidyverse:** Using the tidyverse is like having a skilled personal assistant for R data scientists: it takes the drudgery out of data manipulation and visualization tasks. Two of the most popular packages from the tidyverse ecosystem are {ggplot2} for data visualization and {dplyr} for data manipulation.

