---
title: "How to {ggplot2}"
author: 
- "Carlina Feldmann"
- "Lennart Oelschl&auml;ger"
date: 'Last rendered on `r format(Sys.Date(),"%d.%m.%Y")`'
output: slidy_presentation
---
  
## Why and what 

Welcome to this tiny course on data visualization in R with [{ggplot2}](https://ggplot2.tidyverse.org/)! `r emo::ji("wave")`

### Why do we care?

Potentially, plots can [beautifully inform](https://en.wikipedia.org/wiki/Charles_Joseph_Minard#The_map_of_Napoleon's_Russian_campaign) or [horribly mislead](https://en.wikipedia.org/wiki/Misleading_graph#Truncated_graph). [Colors](https://michaelbach.de/ot) and [shape](http://vis.stanford.edu/files/2010-MTurk-CHI.pdf) matter! `r emo::ji("balance_scale")`

### Why {ggplot2}?

The {ggplot2} package implements a *grammar of graphics*, a series of distinct tasks to make a graphic.

### What is this course about?

Being in decent control of {ggplot2} to produce meaningful plots. 

### What do you need?

Basic R skills + a not-too-old version of R (>= 4.0.0) + RStudio

### At the end of the day...

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library(socviz)
library(maps)
us_states <- map_data("state")
election$region <- tolower(election$state)
us_states_elec <- dplyr::left_join(us_states, election)
ggplot(data = us_states_elec, aes(x = long, y = lat, group = group, fill = party)) +
  geom_polygon(color = "grey", size = 0.1) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  scale_fill_manual(values = c("blue", "red")) +
  theme_void() +
  theme(legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5)) +
  labs(title = "State Winner in 2016", fill = "")
```

### Sources

- Healy (2019): [Data Visualization: A practical introduction](https://socviz.co/)
- Wickham (2016): [ggplot2: elegant graphics for data analysis](https://ggplot2-book.org/)
- Wilkinson (2005): The Grammar of Graphics
- [Tufte](https://en.wikipedia.org/wiki/Edward_Tufte) (1983): The Visual Display of Quantitative Information

### Found mistakes? Have suggestions?

I'm sure you have! Please leave a note [here](https://github.com/loelschlaeger/howtoggplot2/issues/new/choose). `r emo::ji("wish")`

## Our first plot

Load {ggplot2}.

```{r}
# install.packages(ggplot2)
library(ggplot2)
```

We need data, let's go with an excerpt from the famous [Gapminder dataset](https://www.gapminder.org/tools):

```{r}
# install.packages(gapminder)
library(gapminder)
head(gapminder)
str(gapminder)
```

First, we tell the `ggplot()` function what data we use and what variables we wish to see on each axis:

```{r}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) 
```

Something is missing ... `r emo::ji("thinking")` We need an additional layer, a `geom_*` function!

```{r}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) +
  geom_point()
```

There are more of them which we can simply add (literally add!):

```{r, message=FALSE}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p <- p + geom_point() + geom_smooth()
p
```

As a last polishing step for now, we improve the x-axis scale and the plot labels.

```{r, message=FALSE}
p + scale_x_log10(labels = scales::dollar) +
  labs(x = "GDP per capita",
       y = "Life expectancy in years",
       title = "Economic growth as an indicator for life expectancy",
       subtitle = "Data points are country-years",
       caption = "Source: Gapminder")
```

Finally, we can use the `ggsave()` function to save our plot:

```{r, eval = FALSE}
ggsave("some_descriptive_name.pdf")
```

### The {ggplot2} workflow

1. Call `ggplot()`
2. Set `data = ...`
3. Set `mapping = aes(...)`
4. Add one (or more) `geom_*()` functions
5. Adjust the scale and labels

## It's your turn

This course includes tutorials! `r emo::ji("sunglasses")`

Executing the following lines gives you access to the course material:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("loelschlaeger/howtoggplot2")
library(howtoggplot2)
```

To start the tutorial, type:

```{r, eval = FALSE}
practicals()
```

To open a copy of these slides, type:

```{r, eval = FALSE}
slides()
```

To submit an issue on GitHub about this course, type:

```{r, eval = FALSE}
issue()
```

## Facets and more geoms

Our goal is to plot the trajectory of life expectancy over time for each country in the gapminder data. 

```{r}
ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp)) +
  geom_line()
```

We must not forget to group by country! `r emo::ji("light_bulb")`

```{r}
ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp)) +
  geom_line(aes(group = country))
```

But can you make sense of this mess? Luckily, we can group by continents:

```{r}
ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp)) +
  geom_line(aes(group = country)) +
  facet_wrap(~continent)
```

Better don't `facet_wrap(~country)`... Let's polish our plot with the things we already learned:

```{r, message = FALSE}
ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp)) +
  geom_line(color = "grey", aes(group = country)) +
  geom_smooth() +
  facet_wrap(~continent) +
  labs(x = "Year",
       y = "Life expectancy",
       title = "Life expectancy over time on five continents")
```

Notice that we supplied a formula to `facet_wrap`. This can be more advanced, for example (with `facet_grid`):

```{r, message = FALSE, warning = FALSE}
ggplot(data = socviz::gss_sm, mapping = aes(x = age, y = childs)) +
  geom_point(alpha = 0.2) +
  geom_smooth() +
  facet_grid(sex ~ race) +
  labs(x = "Age",
       y = "No. of children",
       title = "Relationship between age and number of children",
       subtitle = "Separated by sex (in rows) and race (in columns)")
```

As a last input for this part, we learn four new geoms.

### Bar plots

```{r}
ggplot(data = socviz::gss_sm, mapping = aes(x = religion)) +
  geom_bar()
```

Using relative instead of absolute counts on the y-axis is covered in the tutorials.

### Histograms

```{r}
ggplot(data = socviz::gss_sm, mapping = aes(x = age)) +
  geom_histogram()
```

We address the message and the warning in the tutorials.

### Density plots

```{r}
library(dplyr)
ggplot(data = filter(gapminder, year == 2007), 
       mapping = aes(x = lifeExp)) +
  geom_density()
```

### Boxplots

```{r}
ggplot(data = filter(gapminder, year == 2007), 
       mapping = aes(x = pop,
                     y = reorder(continent, pop))) +
  geom_boxplot() +
  scale_x_log10() + 
  labs(y = NULL,
       x = "Populations in 2007")
```

We look at a variant on the basic boxplot that {ggplot2} offers in the tutorials.


## Draw Maps

R can work with geographical data, and {ggplot2} can make [choropleth maps](https://en.wikipedia.org/wiki/Choropleth_map).

```{r}
world <- map_data("world")
p <- ggplot(data = world, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "white", color = "black")
plot(p)
```

Instead of the default [Mercator projection](https://en.wikipedia.org/wiki/Mercator_projection), we can use the [Albers projection](https://en.wikipedia.org/wiki/Albers_projection):

```{r}
p + coord_map(projection = "albers", lat0 = 15, lat1 = 45)
```

Now in the tutorials, we will visualize the results of the Trump vs. Clinton election 2016 on a map of the US states.

```





```


```{r}
```


